---
title: "Week 10 Task Set"
author: "Sophia LeBlanc"
format: html
execute:
  echo: true
  warning: true
  message: true
---

#### Q1: Load and Inspect Data ------------------------------------

```{r}
## Load the saved dataset
experiment_data <- readRDS("../data/raw/experiment_data.rds")

## Take a quick look at the first few rows
head(experiment_data)

## Check the structure of the dataset
str(experiment_data)
```

#### Q2: Functions ------------------------------------------------

```{r}
## Generate one or more random numbers between min and max (defaults: one number between 1–10)
getRandomNumber <- function(min = 1, max = 10, number = 1) {
  sample(min:max, number)
}
```

```{r}
## Classify reaction times into "Fast", "Slow", or "Unknown" (default threshold = 500ms)
## Put Code Here:

classify_rt <- function(rt, threshold = 500) {
  if (!is.numeric(rt)) {
    stop("Error: 'rt' must be numeric.")
  }
  if (!is.numeric(threshold) || length(threshold) != 1) {
    stop("Error: 'threshold' must be a single numeric value.")
  }
  return(
    category <- ifelse(
      is.na(rt), "Unknown",
      ifelse(rt < threshold, "Fast", "Slow")
  ))
}

```

#### Q3: Loops vs. Vectorization ---------------------------------------------

```{r}

experiment_data$age <- NA ## new column of NAs

getRandomNumber <- function(min, max) {
  sample(min:max, 1)
}

for(i in 1:nrow(experiment_data)) {
  experiment_data[i, "age"] <-getRandomNumber(18,65)
}

head(experiment_data)

experiment_data$age <- sample(18:65, nrow(experiment_data), replace = TRUE)

## applies sample function to each row of dataset to fill in age column

head(experiment_data)

```

#### Q4: Control Structures on Data Frames ---------------------------

```{r}
for(i in 1:nrow(experiment_data)) {
   experiment_data$rt_category_loop[i] <- classify_rt(experiment_data$rt[i])
}

experiment_data$rt_category_vec <- classify_rt(experiment_data$rt)

## The classify function is already vectorized, as it employs ifelse within. Thanks to AI for helping clarify that logic when I was over-complicating it.

table(loop = experiment_data$rt_category_loop,
      vec  = experiment_data$rt_category_vec,
      useNA = "ifany")

```

#### Q5: Debugging Functions -----------------------------

```{r}
## Create Buggy Function
threshold <- 500

flag_fast <- function(rt_vector) {
  rt_vector < threshold
}

flag_fast(experiment_data$rt)

rm(threshold)

## Above is the code with bug, below is corrected to ensure not reliant on object, threshold is locally defined

flag_fast <- function(rt_vector) {
  threshold <- 500
  rt_vector < threshold
}

```

#### Q6: Debugging Tools in Action -----------------------------

```{r}
subject_means <- tapply(experiment_data$rt,
                        experiment_data$condition,
                        mean)
print(paste("Mean for control:", subject_means["control"]))
print(paste("Mean for incongruent:", subject_means["incongruent"]))

# How many NAs overall?
sum(is.na(experiment_data$rt))

# How many NAs by condition?
tapply(is.na(experiment_data$rt), experiment_data$condition, sum)

# Optional: quick sanity checks
summary(experiment_data$rt)
with(experiment_data, table(condition, is.na(rt)))

## editing code to ignore missing values:

subject_means <- tapply(experiment_data$rt,
                        experiment_data$condition,
                        mean, na.rm = TRUE)
print(paste("Mean for control:", subject_means["control"]))
print(paste("Mean for incongruent:", subject_means["incongruent"]))

```

#### Q7: Using AI for Debugging or Refactoring -----------------------

```{r}
## code with (intentional) bug

## subject_means <- tapply(experiment_data$rt[!is.na(experiment_data$rt)], experiment_data$condition, mean)

## ChatGPT revised, final working code:

subject_means <- tapply(
  experiment_data$rt, 
  experiment_data$condition, 
  mean, 
  na.rm = TRUE
)

```

AI reflection:

- I used ChatGPT. To begin, they did a good job of breaking down the errors and how to fix them into comprehensible steps. They first explained the error itself, in very graspable language for a beginner coder, I think. I like that I first got to understand the logic/reasoning of what was going on (i.e., column length mismatch). ChatGPT even reiterated/summarized main points as it went (e.g., concisely restating the problem). ChatGPT provided revised code and thoroughly explained the changes; their logic aligned with my understanding, like how tapply can apply na.rm = TRUE to rows overall. This explanation was more brief, and I could've asked more questions if I needed to. ChatGPT was also effective at giving alternatives and optional improvements. In one suggestion, I appreciated that they created a NEW cleaned dataset instead of revising the original, which is best practice.

- I changed my code to the "correct approach" after ensuring I understood the changes that were being made. I removed the subsetting and added the na.rm conditional to be applied through tapply. I decided not to implement the data cleaning and the dplyr version as they are unnecessary at this point.

- I executed the revised code to ensure it worked properly. Looking at my global environment, I can see that the value "subject_means" was effectively created, and there are two numbers assigned to it: 511 and 539. There are no NAs seen, so they must have been ignored as intended in the calculation.

#### Q8: Saving and Reloading Objects -----------------------------

```{r}

saveRDS(experiment_data, file = "../data/cleaned/experiment_data.rds")
write.csv(experiment_data, file = "../data/cleaned/experiment_data.csv")

save(experiment_data, subject_means, file = "../data/cleaned/objects.RData")

rm(list = ls())

experiment_data_csv <- read.csv("../data/cleaned/experiment_data.csv")
experiment_data_rds <- readRDS("../data/cleaned/experiment_data.rds")
load("../data/cleaned/objects.RData")

```


